---
title: "COGS202_finalProject"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries
 We are mainly using GAMM4 package to run general additive mixed model, an extension of linear regression.
```{r}
library(psych)
library(plyr)
library(dplyr);
library(lme4);
library(gamm4);
library(lmtest);
library(nlme);
library(ggplot2);
library(gridExtra);
library(gtable);

data <- read.csv('PLING_trimmedData_workingdata.csv')

index = which(complete.cases(data$BEH_WISC_DS_Raw) & complete.cases(data$BEH_VMI_Raw))

raw_data <- data[index,]

```

## Variables of interest

#### Demographic variables
SubjectID: `raw_data$SubjectID`
Age: `raw_data$Age`
Gender: `raw_data$Gender`
Scanner: `raw_data$DeviceSerialNumber.x`

#### Behavioral variables
WISC: `raw_data$BEH_WISC_DS_DSB_Raw`
VMI: `raw_data$BEH_VMI_Raw`

#### DTI variables

Corticospinal tract (CST): DTI_fiber_FA.L_CST, DTI_fiber_FA.R_CST

Anterior thalamic radiations (ATR): DTI_fiber_FA.R_ATR, DTI_fiber_FA.L_ATR 

Corpus callosum (CC): DTI_fiber_FA.CC

Superior longitudinal fasciculus: DTI_fiber_FA.R_SLF, DTI_fiber_FA.L_SLF

Inferior longitudinal fasciculus: DTI_fiber_FA.R_ILF, DTI_fiber_FA.L_ILF

Inferior frontal-occipital fasciculus (IFO): DTI_fiber_FA.R_IFO, DTI_fiber_FA.L_IFO 

Cingulum: CgC (cingulum): DTI_fiber_FA.R_CgC, DTI_fiber_FA.L_CgC, 

test code
```{r}

wisc_dem_fit <- gamm4(BEH_WISC_DS_Raw~ s(Age, k=4, fx=FALSE) + Gender, data= raw_data,  random = ~(1|SubjectID), na.action=na.omit)

summary(wisc_dem_fit$mer)

# adjusted r square
r_sq = summary(wisc_dem_fit$gam)["r.sq"]
plot(wisc_dem_fit$gam)

# AIC
AIC(wisc_dem_fit$mer)

```


## Demographic model

Generating training and testing dataset
```{r}


# training dataset: 80% of the data
smp_size <- floor(0.80 * nrow(raw_data))

MSE_train = rep(NA, 1000)
MSE_test =  rep(NA, 1000)

for (i in 1:1000){
	
set.seed(i)

train_ind <- sample(seq_len(nrow(raw_data)), size = smp_size)

train <- raw_data[train_ind, ]
test <- raw_data[-train_ind, ]

# fitting demographic variables GAMM4 function
wisc_dem_fit_train <- gamm4(BEH_WISC_DS_Raw~ s(Age, k=4, fx=FALSE) + Gender, data= train,  random = ~(1|SubjectID), na.action=na.omit)

#summary(wisc_dem_fit_train$gam)

MSE_train[i] <- mean((train$BEH_WISC_DS_Raw - fitted(wisc_dem_fit_train$gam))^2)

# using the estimated parameter estimates to predict the test data
pred_test <- predict(wisc_dem_fit_train$gam, test, na.action=na.exclude)

MSE_test[i] <- mean((test$BEH_WISC_DS_Raw - pred_test)^2)

}

Mean_MSE_train = mean(MSE_train)
Mean_MSE_test = mean(MSE_test)

Mean_MSE_train
Mean_MSE_test
```

## DTI model




